Let There Be Light: \\ Early Observations of the Normal Type Ia Supernova \abc\ \\Show No Evidence for a Dark Period

### 2017-07-31 doc/ms.tex
Update introduction to reflect the new conclusions regarding diffuse interaction

Early phase observations of SNe Ia from a white dwarf (WD)$+$non-degenerate binary
may detect excess emission, relative to most type Ia SNe, due to the 
collision of the SN ejecta with the non-degenerate companion 
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess 
emission was first detected in iPTF\,14atg  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as a
SN ejecta-companion collision. While such emission requires a 
favorable geometric alignment and is only expected in $\lesssim$ 
10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have 
searched for signatures of an ejecta-companion interaction, 
typically resulting in non-detections 
(e.g., \citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}). A possible exception is SN\,2012cg, 
which exhibited excess blue emission in its early-phase light curve 
\citet{2016ApJ...820...92M} (though see 
\citealt{2016arXiv161007601S} for an interpretation that does not 
invoke ejecta-companion interaction). 
% With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
% signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

The vast majority of SNe Ia are observed 
to be powered purely by the radioactive decay of $^{56}$Ni. 
While the detection of SN shock cooling or ejecta-companion interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can fundamentally alter the appearance of the SN shortly after explosion. SNe Ia experience a dark period after
the SN shock breakout but before radioactive energy  
diffuses into the photosphere \citep{2014ApJ...784...85P}. The 
duration of this dark period is set by how the newly synthesized 
$^{56}$Ni is mixed and deposited into different layers of the ejecta.
In the case of strong mixing, the dark period is short, or 
non-existent, as photons from $^{56}$Ni decay can reach the 
photosphere rapidly. This also leads to blue optical colors and a 
rapid initial rise in the observed light curve. If the mixing is 
weak and the $^{56}$Ni is confined to the innermost layers of the 
ejecta, the dark period can last for several days as the radioactive 
energy diffuses to the photosphere. The early evolution of such SNe 
results in redder colors and a more moderate rise in luminosity 
\citep{2016ApJ...826...96P}. Thus, the early light curves of even non-exotic type Ia SNe convey information about their progenitor systems by constraining the distribution of synthesized $^{56}$Ni, which in turn constrains the explosion mechanism.

In this paper, we report observations of an extraordinarily young SN
Ia, \abc, which was discovered by the intermediate Palomar
Transient Factory (iPTF) on 2016 April $3.36$\footnote{All times in this
  paper are in UTC.} at $\textrm{R.A.}=13^h34^m45.49^s$,
$\textrm{Dec.}=+13^d51^m14.3^s$ (J2000) with a $g_\mathrm{PTF}$-band magnitude of
$21.31\pm0.27$ \citep{2016ATel.8907....1M}. The
transient is spatially coincident with a tidal tail of the galaxy
NGC\,5221, which lies at a distance of $\sim$100\,Mpc. \abc\ is not detected to a limit of $g=22.1$\,mag on April $2.42$, less than 1 d prior to discovery, and rose by $\sim$2 mag in the 24 hr following its initial detection. Our spectroscopic follow-up
campaign classified \abc\ as a normal SN Ia
\citep{2016ATel.8909....1C}. Our observations and analysis provide multiple lines of evidence that the $^{56}$Ni produced by \abc\ was strongly mixed throughout the ejecta.

>>>>

Early phase observations of SNe Ia from a white dwarf (WD)$+$non-degenerate binary
may detect excess emission, relative to most type Ia SNe, due to the 
collision of the SN ejecta with the non-degenerate companion 
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess 
emission was first detected in iPTF\,14atg  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as a
SN ejecta-companion collision. While such emission requires a 
favorable geometric alignment and is only expected in $\lesssim$ 
10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have 
searched for signatures of an ejecta-companion interaction, 
typically resulting in non-detections 
(e.g., \citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}). Possible exceptions include SN\,2012cg, 
which exhibited excess blue emission in its early-phase light curve 
(\citealt{2016ApJ...820...92M}; though see 
\citealt{2016arXiv161007601S} for an interpretation that does not 
invoke ejecta-companion interaction), and SN\,2017cbv, which shows a clearly resolved ``bump'' in the early $UBg$ light curves \citep{2017arXiv170608990H}. 
% With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
% signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

Interaction is not limited to systems with a non-degenerate companion, however, as WDs enshoruded in diffuse material following a binary merger or expanded due to a pre-explosion pulsation can give rise to ejecta-interaction signatures (see e.g., \citealt{2014MNRAS.441..532D}). Models of this scenario naturally produce \ion{C}{2} absorption that is comparable in strength to \ion{Si}{2} in the days after explosion \citep{2014MNRAS.441..532D}, as was observed in SN\;2013dy \citep{2013ApJ...778L..15Z} and SN\;2017cbv \citep{2017arXiv170608990H}.

The vast majority of SNe Ia are observed 
to be powered purely by the radioactive decay of $^{56}$Ni. 
While the detection of SN shock cooling or ejecta interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can fundamentally alter the appearance of the SN shortly after explosion. SNe Ia experience a dark period after
the SN shock breakout but before radioactive energy  
diffuses into the photosphere \citep{2014ApJ...784...85P}. The 
duration of this dark period is set by how the newly synthesized 
$^{56}$Ni is mixed and deposited into different layers of the ejecta. In the case of strong mixing, the dark period is short, or 
non-existent, as photons from $^{56}$Ni decay can reach the 
photosphere rapidly. This also leads to blue optical colors and a 
rapid initial rise in the observed light curve. If the mixing is 
weak and the $^{56}$Ni is confined to the innermost layers of the 
ejecta, the dark period can last for several days as the radioactive energy diffuses to the photosphere. The early evolution of such SNe results in redder colors and a more moderate rise in luminosity \citep{2016ApJ...826...96P}. Thus, the early light curves of even non-exotic type Ia SNe convey information about their progenitor systems by constraining the distribution of synthesized $^{56}$Ni, which in turn constrains the explosion mechanism.

In this paper, we report observations of an extraordinarily young SN Ia, \abc, which was discovered by the intermediate Palomar Transient Factory (iPTF) on 2016 April $3.36$\footnote{All times in this paper are in UTC.} at $\textrm{R.A.}=13^h34^m45.49^s$, $\textrm{Dec.}=+13^d51^m14.3^s$ (J2000) with a $g_\mathrm{PTF}$-band magnitude of $21.31\pm0.27$ \citep{2016ATel.8907....1M}. The transient is spatially coincident with a tidal tail of the galaxy NGC\,5221, which lies at a distance of $\sim$100\,Mpc. \abc\ is not detected to a limit of $g=22.1$\,mag on April $2.42$, less than 1 d prior to discovery, and rose by $\sim$2 mag in the 24 hr following its initial detection. Our spectroscopic follow-up
campaign classified \abc\ as a normal SN Ia
\citep{2016ATel.8909....1C}. Our observations and analysis show that the early evolution of \abc\ exhibited several distinct properties relative to SN\,2011fe. We interpret those differences as arising from either strong $^{56}$Ni mixing or ejecta interaction with diffuse material, or a combination of the two.
